Dispersal Through Seasonal Landscapes

### 2021-06-24 Chapter_2/04_Manuscript/Manuscript.tex
various updates

%------------------------------------------------------------------------------
%	Titlepage: Header
%------------------------------------------------------------------------------
\title{Step by Step: Using Step Selection Functions to Simulate Dispersal and
Assess Landscape Connectivity}

@@

%------------------------------------------------------------------------------
%	Abstract
%------------------------------------------------------------------------------
\newpage
\begin{abstract}
Dispersal is an important process that allows species to avoid inbreeding, to
colonize new habitats and to reinforce non-viable subpopulations. Successful
dispersal thus represents a crucial pre-requisite for long-term species
persistence in wild animal populations. However, the ability to disperse is
contingent a sufficient degree of landscape connectivity, which is why the
estimation of connectivity and identification and preservation of dispersal
corridors has become a task of extraordinary importance for conservation
authorities worldwide.

Over the past two decades, ecologists have primarily relied on analytical tools
such as least-cost analysis and circuit theory to model and investigate
landscape connectivity. Despite their usefulness for a diverse suite of
ecological applications, both methods make several restricting assumptions that
limit their suitability in reality. To address these shortcomings, dispersal
simulations from individual-based movement models have been proposed and
applied. Yet, due to the almost infinite number of non-trivial decisions a
modeler faces when parametrizing such models, a unified and objective framework
is missing.

Recent innovations in movement ecology have brought forward novel opportunities
to study animal dispersal and estimate landscape connectivity. In particular,
the rich suite of resource selection functions, namely point-, step-, and
path-selection functions, have undergone substantial improvements over the past
years. Most notably, step-selection functions have been generalized to
\textit{integrated} step selection functions, which essentially represent fully
mechanistic movement models based on which an individual's movement could be
simulated. While such models have been applied to study \textit{steady-state}
utilization distribution resident animals, a similar approach may be useful for
investigating \textit{transient} movement behavior and study landscape
connectivity.

Here, we showcase the use of integrated step selection functions as a simple
individual-based and spatially explicit model to simulate dispersal of the
endangered African wild dog across the world's largest transboundary
conservation area, the Kavango-Zambezi Transfrontier Conservation Area
(KAZA-TFCA). For this, we utilize data collected on 16 dispersing wild dog
coalitions in combination with relevant habitat covariates. We analyse the data
using integrated step selection functions and parametrize a fully mechanistic
model describing how dispersing wild dogs move through the landscape. Based on
this model, we simulate 80'000 dispersers moving across the extent of the
KAZA-TFCA. We then generate a set of maps, each focused on a different aspect of
connectivity within the KAZA-TFCA. Finally, we discuss the benefits and pitfalls
of such a simulation-based approach and highlight potential improvements to be
made in the future.

\end{abstract}

@@

% Importance of Dispersal & Connectivity
\subsection{Importance of Dispersal \& Connectivity (90\%)}
Dispersal is defined as the movement of individuals away from their natal
location to the site of first reproduction \cite{Howard.1960}. It is a vital
process governing the social structure of wild animal populations that are
distributed in space \citep{Hanski.1998, Clobert.2012} and may strongly affect
population dynamics at different spatial and social scales \citep{Hanski.1999a,
Clobert.2012}. Dispersal allows species to avoid inbreeding and maintain genetic
diversity \citep{Perrin.1999, Perrin.2000, Frankham.2002, Leigh.2012,
Baguette.2013}, to rescue small, non-viable populations \citep{Brown.1977}, and
to promote the colonization of unoccupied habitats \citep{Hanski.1999b,
MacArthur.2001}. However, successful dispersal requires a sufficient degree of
landscape connectivity \citep{Fahrig.2003, Clobert.2012}, which is why the
identification and protection of major dispersal corridors has become a
fundamental task in conservation science \citep{Nathan.2008, Doerr.2011,
Rudnick.2012}. Importantly, the ability to pinpoint relevant dispersal hotspots
requires information on movement behavior during dispersal and knowledge about
factors that limit dispersal and therefore connectivity \citep{Baguette.2013,
Vasudev.2015}.

% Advancements in GPS Technology & Movement Ecology
\subsection{Advancements in GPS Technology \& Movement Ecology (90\%)}
Thanks to novel technologies developed over the past decades, particularly of
GPS/Satellite radio-collars, the use of GPS data to study animal dispersal and
connectivity has accelerated \citep{Elliot.2014, Jonsson.2016, Williams.2019}.
Additionally, the advent of publicly accessible satellite imagery and
sophisticated remote sensing techniques to represent the physical landscape
through which individuals disperse have heralded a ``golden age of animal
tracking'' \citep{Kays.2015}. Concurrently, the availability of large amounts of
empirical data and an increased computational power have led to the development
of numerous techniques to study dispersal and highlight critical corridors
between subpopulations \citep{Boyce.2002, Fortin.2005, Cushman.2010,
Zeller.2012, Diniz.2020}.

% Resource Selection & Connectivity
\subsection{Resource Selection \& Connectivity (90\%)}
\textit{Resource selection functions} \citep{Boyce.2002} and derived methods
such as \textit{step selection functions} \citep{Fortin.2005} and \textit{path
selection functions} \citep{Cushman.2010} have proven particularly useful for
studying animal movement \citep{Fieberg.2020} and modeling connectivity
\citep{Diniz.2020}. These methods allow estimating habitat preferences of the
focal species by comparing covariates at locations visited by the animal to the
same covariates at locations available to, but not visited by the animal
\citep{Boyce.2002, Fortin.2005, Cushman.2010, Thurfjell.2014}. The so estimated
preferences can then be used to predict a permeability surface, indicating the
expected ease at which an animal can traverse a given area \citep{Spear.2010,
Zeller.2012, Etherington.2016}. Ultimately, the permeability surface serves as
input to a connectivity model that is used to reveal movement corridors
\citep{Diniz.2020}. In this regard, two of the most prominent connectivity
models are least-cost path analysis (LCP analysis; \citealp{Adriaensen.2003})
and circuit theory (CT \citealp{McRae.2006, McRae.2008}), both graph-based
methods that estimate conductance of the landscape. Despite their intuitive
nature and ease of use, both methods make rigorous assumptions about animal
movement that are often not fulfilled in reality \citep{Diniz.2020}.

% Issues with Least-Cost Paths & Circuit Theory
\subsection{Issues with Least-Cost Paths (90\%)}
In LCP analysis, for instance, a least costly path always exists, even if
associated movement costs are unreasonably high and will never be incurred by a
dispersing individual. The method also presumes that animals have an infinite
perceptual range and a preconceived end-point in mind, such that they choose a
cost-minimizing route accordingly. These assumptions may be reasonable for
migrating animals, yet they are unlikely to hold true for dispersers, which
typically move over long distances into unknown territory \citep{Koen.2014,
Abrahms.2017, Cozzi.2020}. Another issue of LCPs analysis concerns the fact that
least-costly routes, by their very nature, are only one pixel wide. This implies
that the width of inferred movement routes depends on the resolution of chosen
covariate layers and may not be biologically meaningful \citep{Diniz.2020}.
Although some of these deficiencies can be addressed using less stringent
versions of the LCP algorithm (e.g. least-cost \textit{corridors}
\citep{Pinto.2009}, \textit{thresholded} least-cost paths \citep{Landguth.2012},
and \textit{randomized} least-cost paths \citep{Panzacchi.2016,
VanMoorter.2021}), a certain degree of arbitrariness remains.

\subsection{Issues with Circuit Theory (90\%)}
CT entails similarly unreasonable restrictions that are hardly ever met. Because
CT only allows movements from a source cell to its 4 or 8 adjacent cells, it
implicitly posits a perceptual range of a single pixel. Assuming that covariate
layers are usually resolve between 30 m and 1 km, this hardly ever renders the
true capability of animals to perceive the environment. Moreover, CT builds on
the assumption of a complete random walk \citep{Diniz.2020}, implying that
directional biases cannot be rendered. Nevertheless, directionality is a
commonly observed characteristic in animal movement \citep{Bovet.1991,
Schultz.2001}, especially in dispersing individuals \citep{Cozzi.2020,
Hofmann.2021}.

\subsection{Issues of Both Methods}
Finally, neither LCP analysis nor CT are capable of rendering the temporal
dimension of dispersal. Statements about the expected duration required to
traverse a certain corridor are therefore impossible. Likewise, because movement
is not modeled explicitly, neither of the methods allows to render interactions
between movement and habitat preferences of the focal species. Connectivity
therefore merely arises in result to the landscape structure, which is usually
referred to as structural connectivity. While structural connectivity yields
insights in the \textit{potential} of the landscape to be traversed, it does not
enable to quantify the \textit{actual} gene flow through the area. Consequently,
a functional on connectivity, which also renders the behavioral response of the
animal with respect to prevailing habitat conditions, is often more desirable
\citep{Tischendorf.2000, Baguette.2013}.

% What about IBMMs?
\subsection{What about IBMMs? (90\%)}
To address the issues inherent to LCPs and CT, individual-based movement models
(IBMMs) have been proposed and applied \citep{Diniz.2020}. In these models,
dispersal is simulated spatially explicitly, based on movement rules that
determine how individuals move over and interact with the prevailing landscape
\citep{Gustafson.1996, Gardner.2004, Graf.2007, KramerSchadt.2004, Revilla.2004,
Revilla.2008, Kanagaraj.2013, Clark.2015, Allen.2016, Hauenstein.2019,
Zeller.2020, Vasudev.2021}. Using the simulated trajectories, one can calculate
a set of connectivity metrics, such as inter-patch connectivity and traversal
frequency to reveal major dispersal corridors \citep{Kanagaraj.2013,
BastilleRousseau.2018, Hauenstein.2019, Zeller.2020}. However, while IBMMS can
be employed to overcome any of the shortcomings intrinsic to LCPs and CT and
provide a more functional view on connectivity, they can be challenging to fit
and require vast amounts of data collected during dispersal \citep{Diniz.2020}.
Moreover, most individual-based models are heavily influenced by subjective
decisions, rendering replication and validation of results difficult.

% Step Selection Analysis
\subsection{Step Selection Analysis (90\%)}
Here, we investigate the usefulness of integrated step selection functions
(ISSFs, \citealp{Avgar.2016}), as a relatively simple but powerful IBMM based on
which dispersal can be simulated. While regular SSFs were intended to learn
about relative habitat preferences of the focal species \citep{Fortin.2005,
Thurfjell.2014, Avgar.2017}, the method has been generalized and now enables to
jointly study habitat and movement preferences, as well as potential
interactions between movement and habitat preferences \citep{Avgar.2016,
Signer.2017, Fieberg.2020}. ISSFs therefore provide a relatively simple method
to model complex movement behavior where movement results from two intertwined
behavioral kernels (e.g. \citealp{Prokopenko.2017, Munden.2020}). Importantly, a
parametrized ISSF model can be viewed as a fully mechanistic movement model
based on which individual movement trajectories can be simulated
\citep{Avgar.2016, Signer.2017}. In fact, \cite{Signer.2017} used ISSF to
simulate steady state utilization distributions of resident animals. However,
the degree to which such simulations are helpful in detecting movement corridors
and modeling landscape connectivity is unknown.

% Study Species & Study Area
\subsection{Study Species \& Study Area (90\%)}
One of the species for which long-term viability relies on sufficient landscape
connectivity is the endangered African wild dog \textit{Lycon pictus}. While
once present across entire sub-Saharan Africa, wild dogs have disappeared from a
vast majority of their historic range due to persecution by humans, habitat
fragmentation and destruction, and deadly diseases \citep{Woodroffe.2012}. As of
today, only 6'000 free-ranging individuals remain in small and spatially
scattered subpopulations \citep{Woodroffe.2012}. Within those subpopulations,
wild dogs form cohesive packs comprising 8 to 12 adults and their offspring
\cite{McNutt.1995}. After reaching sexual maturity, male and female offspring
form same-sex coalitions and disperse from their natal pack \citep{McNutt.1996,
Behr.2020}. New packs are formed when dispersing coalitions join unrelated
opposite-sex dispersing coalitions \citep{McNutt.1996}. Dispersing wild dogs can
cover several hundred kilometers across a variety of landscapes
\citep{DaviesMostert.2012, Masenga.2016, Cozzi.2020, Hofmann.2021}. One of the
few strongholds for this species lies near the Moremi Game Reserve in northern
Botswana, which is part of the world's largest transboundary protected area,
namely the Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA). This
area has originally been intended to facilitate migration of elephants, but is
expected to benefit a multitude of other species \citep{Elliot.2014,
Brennan.2020, Hofmann.2021}.

% Previous Paper
\subsection{Previous Paper (90\%)}
In a previous study, we assessed landscape connectivity for dispersing African
wild dogs within the KAZA-TFCA using a least-cost methods \citep{Hofmann.2021}.
Specifically, we fitted a basic habitat selection model using which we predicted
a permeability surface and computed least-cost paths and corridors. We now
expand this model and use ISSFs to develop a more detailed movement model of
dispersing wild dogs (\Cref{GraphicalAbstract}). We then employ the model to
simulate dispersers moving across the KAZA-TFCA. Based on simulations, we
generate three distinc connectivity maps, all sheding light onto a different
aspect of connectivity. With this work we hope to show that a simulation-based
approach yields several benefits over traditional connectivity modeling
techniques such as least-cost analyses and circuit theory. Most importantly,
simulations provide a more generic view on how connectivity emerges and to which
degree connectivity depends on the dispersal duration. In addtion, by generating
proper dispersal trajectories, network theory can be applied to calculate
network metrics that are pertinent to connectivity analysis.

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width = \textwidth]{99_GraphicalAbstract.pdf}
    \caption{Flowchart of the simulation-based connectivity analysis. First, GPS
    data and habitat covariates have to be collected. The combined data is then
    analyzed in an integrated step selection model, which enables the
    parametrization of the focal species' habitat and movement kernels. The
    parametrized model is then treated as an individual-based movement model and
    used to simulate dispersal trajectories. Ultimately, simulated trajectories
    serve to produce a set of maps that are pertinent to landscape connectivity.
    This includes a heatmap, indicating the relative traversal frequency across
    the study area, a betweenness map, highlighting movement corridors and
    bottlenecks, and, finally, an inter-patch connectivity map, where the
    frequency of connections and their average duration can be depicted. Photos:
    Whom to cite? Vectronics or Photographers?}
    \label{GraphicalAbstract}
  \end{center}
\end{figure}

\section{Methods}
\subsection{Study Area (90\%)}
The study area was defined by a bounding box centered at -17\degree 13'9''S,
23\degree 56'4''E (\Cref{StudyArea}a) stretching over 1.3 Mio.
km\textsuperscript{2} and ecompassing the entire KAZA-TFCA (\Cref{StudyArea}b).
The KAZA-TFCA represents the world's largest transboundary conservation area and
comprises parts of Angola, Botswana, Namibia, Zimbabwe, and Zambia. It covers a
total of 520'000 km\textsuperscript{2} and hosts a diverse landscape, ranging
from savanna to grassland and from dry to moist woodland habitats. In its center
lies the Okavango Delta, a dominant hydrogeographical feature and the world's
largest flood-pulsing inland delta. The wet season within the KAZA-TFCA lasts
from November to March and is out of phase with the flood in the Okavango Delta,
which peaks between July and August \citep{McNutt.1996, Wolski.2017}. Although
large portions within the KAZA-TFCA are designated national parks or other
protected areas, considerable human influence remains due to roads, agricultural
sites and settlements and villages that are distributed across the KAZA-TFCA's
landscape.

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width = \textwidth]{99_StudyArea.png}
    \caption{Illustration of the study area located in southern Africa. (a) The
    study area was confined by a bounding box spanning the entire KAZA-TFCA and
    encompassing parts of Angola, Namibia, Botswana, Zimbabwe, and Zambia. (b)
    The KAZA-TFCA represents the world's largest terrestrial conservation area
    and covers a total of 520'000 km\textsuperscript{2}. Its purpose is to
    re-establish connectivity between already-existing national parks (dark
    green) and other protected areas (light green). The dispersal data used in
    this study was collected on a free-ranging African wild dog population
    inhabiting the Moremi National Park in northern Botswana.}
    \label{StudyArea}
  \end{center}
\end{figure}

\subsection{GPS Relocation Data (90\%)}
Between 2011 and 2019, we collected GPS relocation data on dispersing wild dogs
from a free-ranging wild dog population inhabiting the Moremi National Park in
northern Botswana \citep{Cozzi.2020, Hofmann.2021}. We selected potential
dispersers based on age, pack size, number of same‐sex siblings within the pack,
and presence of unrelated opposite-sex individuals in the pack
\citep{McNutt.1996, Behr.2020}. We immobilized selected individuals using a
cocktail of Ketamine/Xylazine/Atropine \citep{Osofsky.1996, Cozzi.2020} that was
injected by dart, fired from a CO\textsubscript{2}-pressurized gun
(\textit{DAN-Inject, Denmark}). Immobilized individuals were fitted with
GPS/Satellite radio collars (\textit{Vertex Lite; Vectronic Aerospace GmbH,
Berlin}) that guaranteed automated drop-off through a decomposable piece of
cotton. Handling and collaring of all individuals was supervised by a
Botswana-registered wildlife veterinarian and all individuals quickly rejoined
their pack after immobilization.

16 collared individuals eventually dispersed, each in a separate same-sex
dispersal coalition (7 female and 9 male coalitions). During dispersal, collars
were programmed to record a GPS fix every 4 hours. Collected relocations were
regularly transmitted over the Iridium satellite system, which allowed remote
tracking of individuals, even if they left the main study area and crossed
international borders. Because behavior during dispersal is more pertinent for
assessing landscape connectivity \citep{Elliot.2014, Abrahms.2017}, we discarded
all data that was collected during residency and only retained GPS data recorded
during dispersal. In some instances, exact dispersal dates were known from field
observations. Where this was not the case, determined dispersal phases using the
net-squared displacement metric. Net squared displacement measures the squared
Euclidean distance of a GPS relocation to a reference point \citep{Borger.2012},
which in our case was set to the center of each individual's natal home range.
As such, dispersal was deemed to have started when an individual left its natal
home range and ended once individuals became sedentary again. As previous
research revealed similar behavior of females and males during dispersal
\citep{Woodroffe.2019, Cozzi.2020}, we did not distinguish between sexes. After
collection, we converted collected GPS coordinates (n = 4'169) to steps, where
each step represented the straight-line distance traveled by and individual
between two consecutive GPS relocations \citep{Turchin.1998}. To ensure a
regular sampling interval, we removed fixes that were not successfully collected
on the 4-hourly schedule (\( \pm \) 15 minutes).

\subsection{Covariates (90\%)}
We represented the physical landscape across the study area using a set of
habitat covariates that included water-cover, distance to water, woodland-cover,
and shrub/grassland-cover. Because water cover greatly changes within and
between years in the Okavango Delta, we applied a remote sensing algorithm and
generated frequently updated water cover layers and corresponding distance to
water layers (see \citealp{Wolski.2017} and Appendix A3 in
\citealp{Hofmann.2021}). Resulting water layers thus temporally aligned with our
dispersal events. We furthermore computed a proxy for human influence, rendering
anthropogenic pressures stemming from human-density, agricultural sites, and
roads. All spatial layers were coarsened or interpolated to a target resolution
of 250 m by 250 m. Further details on the sources and preparation of each
habitat covariate are given in \cite{Hofmann.2021}.

Besides habitat covariates, we computed movement metrics that we used as
movement covariates in our models. Movement metrics were calculated for each
step and included the step length (\textsf{sl}), its natural logarithm
(\textsf{log(sl)}), and the cosine of the relative turning angle
(\textsf{cos(ta)}) (for details see \citep{Avgar.2016, Fieberg.2020}). Because
wild dogs follow a diurnal activity pattern, we also coded a binary variable
(\textsf{LowActivity}) indicating whether a step was realized during periods of
low wild dog activity (17:00 to 09:00 local time) or high wild dog activity
(09:00 to 17:00 local time). Handling and manipulation of all data, as well as
all models and simulations were implemented with the statistical software {\tt
R}, version 3.6.6 \citep{R.2019}. Several helper functions were written in {\tt
C++} and imported into {\tt R} using the {\tt Rcpp} package
\citep{Eddelbuettel.2011, Eddelbuettel.2013}

\subsection{Movement Model (80\%)}
We used ISSFs to parametrize a mechanistic movement model for dispersing wild
dogs \citep{Avgar.2016}. To conduct ISSF analysis, we paired each realized step
with 24 random steps. An observed and its 24 random steps formed a stratum and
received a unique identifier. We generated random steps by sampling random
turning angles from a uniform distribution (\(-\pi, +\pi\)) and step lengths
from a gamma distirbution that was fitted to realized steps (scale = 6'308,
shape = 0.37). Along each step, we extracted and averaged spatial covariates
using the {\tt velox} package \citep{Hunziker.2021}. We also calculated the
movement metrics \textsf{sl}, \textsf{log(sl)}, and \textsf{cos(ta)} for each
observed and random step. To facilitate model convergence, we standardized all
continuous covariates to a mean of zero and a standard deviation of one. Since
correlation among covariates was low (\(|r| > 0.6\); \citealp{Latham.2011}), we
retained all of them for modeling.

To contrast realized steps (scored 1) and random steps (scored 0), we assumed
that animals assigned a selection score \(w(x)\) of the exponential form to each
step \citep{Fortin.2005}. The selection score \(w(x)\) of each step depended on
its associated covariates (\(x_1, x_2, ..., x_n\)) and on the animal's
preferences (i.e. relative selection strengths; \citealp{Avgar.2017}) towards
these covariates (\(\beta_1, \beta_2, ..., \beta_n\)):

\begin{equation}
\label{EQ1}
  w(x) = exp(\beta_1 x_1 + \beta_2 x_2 + ... + \beta_n x_n)
\end{equation}

The probability of a step being realized was then contingent on the step's
selection score, as well as on the selection scores of all other step in the
same stratum:

\begin{equation}
\label{EQ2}
  P(Y_{i} = 1 | Y_{1} + Y_{2} + ... + Y_{i} = 1) =
  \frac{w(x_{i})}{w(x_{1}) + w(x_{2}) + ... + w(x_{i})}
\end{equation}

To estimate the preferences of interest, we ran conditional logistic regression
analysis in the r-package {\tt glmmTMB}. To handle multiple individuals, we
applied the mixed effects technnique developed by \citep{Muff.2020}, which
allows to effectively model random slopes. Thus, we treated animal IDs as random
effect and modeled random slopes for each covariate. We fixed the random
intercept variance at an arbitrary high value of 10\textsuperscript{6} to make
use of the  ``poission''-trick \citep{Muff.2020}.

The formula for the movement model was based on the habitat selection model for
dispersing wild dogs presented in \cite{Hofmann.2021}. In the original model, no
interactions between habitat covariates (\textsf{Water,
DistanceToWater\textsuperscript{0.5}, Woodland, Shrubs/Grassland, Human
Influence}) and movement covariates (\textsf{sl, log(sl), cos(ta)}) were
considered. Here, we slightly expanded this base model and proposed interactions
between all movement and habitat covariates. More specifically, we started with
the base model and incrementally increased model complexity by adding all
possible two-way interactions between habitat covariates and movement
covariates. For instance, for the covariate \textsf{Water}, we proposed the
interactions \textsf{Water:log(sl)}, \textsf{Water:log(sl)}, and
\textsf{Water:cos(ta)}. Besides those combinations, we also proposed the
interactions \textsf{sl:cos(ta)} and \textsf{log(sl):cos(ta)} to account for a
correlation between turning angles and step lengths, as well as the interactions
\textsf{sl:LowActivity} and \textsf{log(sl):LowActivity} to account for the fact
that step lengths may differ due to wild dogs' diurnal activity pattern. To
compare competing models and assess the most parsimonious movement model, we ran
stepwise forward model selection based on Akaike's Information Criterion (AIC,
\citealp{Burnham.2002}).

We validated the predictive power of the most parsimonious movement model using
k-fold cross-validation for case-control studies as suggested by
\cite{Fortin.2009}. For this, we randomly assigned 80\% of the strata to a
training set and the remaining 20\% to a testing set. Using the training data we
parametrized a movement model based on which we predicted selection scores
\(w(x)\) for all steps in the test data. Within each stratum, we then assigned
ranks 1-25 to each step based on predicted selection scores, where rank 1 was
given to the step with the highest score \(w(x)\). Across all strata we
determined the realized step's rank and we calculated rank frequencies of
realized steps across all strata. Finally, we computed Spearman's rank
correlation between ranks and associated frequencies \(r_{s, realized}\). We
replicated the entire procedure 100 times and computed the mean correlation
coefficient (\(\bar{r}_{s, realized}\)), as well as its 95\% confidence interval
across all replicates. For comparison, we repeated the same procedure 100 times
assuming random preferences, which we implemented by discarding the realized
step from all strata and identifying the rank of a random step in each stratum.
Again, we calculated Spearman's rank correlation coefficient (\(r_{s,
random}\)), its mean across repetitions (\(\bar{r}_{s, random}\)), and its 95\%
confidence interval. This validation ultimately proofs a significant prediction
in case the confidence intervals of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s,
random}\) do not overlap.

\subsection{Dispersal Simulation (80\%)}
We used the most parsimonious movement model to simulate 80'000 virtual
dispersers moving across the KAZA-TFCA. The simulation resembled an inverted
ISSF and was set up as follows. (1) We defined a random source point and assumed
a random initial orientation of the animal. (2) Departing from the source point,
we generated 25 random steps by sampling turning angles from a uniform
distribution (\(-\pi, +\pi\)) and step lengths from our fitted gamma
distribution. Similar to the input data, each random step represented the
straight line movement within 4 hours. To prevent unreasonably large steps, we
capped sampled step lengths to a maximum of 35 km, which corresponded to the
farthest distance ever traveled within 4 hours by one of our dispersers. (3)
Along each random step, we extracted and averaged habitat covariates and we
calculated movement covariates. To ensure compatible scales, we standardized
extracted values using the same parameters applied to our input data. (4) We
applied the parameterized movement model to predict the selection score
\(w(x)\), which we translated into probabilities using \Cref{EQ2}. (5) We
sampled one of the random steps based on predicted probabilities and determined
the animal's new position. We repeated steps (2) to (5) until 2'000 steps were
realized, implying a total 160 Mio. simulated steps.

To minimize the influence of edge effects and to deal with random steps leaving
the study area, we followed \citep{Koen.2010} and artificially expanded all
covariate layers by adding a 100 km wide buffer zone. Within the buffer zone, we
randomized covariate values by resampling values from the original covariate
layers. Through this buffer zone, simulated dispersers were able to leave and
re-enter the main study area. In cases where proposed random steps transgressed
the border of this buffer zone, we resampled transgressing steps until they
fully lied within the buffer, thereby forcing simulated individuals to ``bounce
off'' from such invisible borders.

\subsection{Source Points (90\%)}
We released 80'000 virtual dispersers from 80'000 unique source points
distributed across the study area. 50'000 virtual dispersers were released from
randomly selected source points within contiguous protected areas larger \(>\)
700 km\textsuperscript{2} (\Cref{SourcePoints}a), which conforms to average home
range requirements of resident wild dogs \citep{Pomilia.2015} and allowed us to
remove patches too small to host viable populations. By distributing source
points randomly, the number of source points per km\textsuperscript{2} was
approximately equal within protected areas. To render potential immigrants into
the study system, we released another 30'000 dispersers at random locations
inside the 100 km wide buffer zone surrounding the main study area
(\Cref{SourcePoints}b).

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width = \textwidth]{99_SourceAreas.png}
    \caption{(a) Different source areas from which we released virtual
    dispersers. We only considered contiguous protected areas (national parks,
    game reserves, and forest reserves) that were larger than 700
    km\textsuperscript{2} (green). This area corresponds to the average home
    range requirement for viable wild dog populations \citep{Pomilia.2015}. To
    render potential immigrants into the study system, we also initiated
    dispersers within a buffer zone (blue) surrounding the main study area. (b)
    Source points from which dipsersers were released. 50'000 dispersers were
    released from the main study area (green dots) and another 30'000 dispersers
    within the virtual buffer (blue dots).}
    \label{SourcePoints}
  \end{center}
\end{figure}

\subsection{Convegence (80\%)}
To verify that the number of simulated individuals sufficed to ensure reliable
estimates of connectivity, we evaluated how the relative traversal frequency
across the landscape depended on the number of simulated trajectories.
Specifically, we placed 1'000 ``checkpoints'' each with an extent of 5 km x 5 km
at random locations inside the main study area. We then determined the relative
traversal frequency by simulated trajectories through each checkpoint for
different numbers of simulations (1 to 50'000). We replicated each random sample
100 times and computed the mean across replicates, as well as the 95\%
xx-interval. Note that we only considered the 50'000 trajectories originating
from the main study area.

\subsection{Heatmap (100\%)}
To identify dispersal hotspots across our study area, we created a heatmap
indicating the absolute frequency at which each raster-cell in the study area
was visited by virtual dispersers \citep{Hauenstein.2019, Peer.2008}. For this,
we rasterized all simulated trajectories and tallied them into a single map. If
the same trajectory crossed a raster-cell twice, it was only counted once,
thereby mitigating potential biases resulting from individuals that were trapped
and moved in circles. To achieve high performance rasterization, we used the
R-package {\tt terra} \citep{Hijmans.2020}.

\subsection{Betweenness (80\%)}
To pinpoint areas of exceptional relevance for connecting remote regions inside
our study area, we converted simulated trajectories into a network and
calculated betweenness scores \citep{BastilleRousseau.2018}. For this, we
overlaid the study area (including the buffer) with a regular raster resolved at
5 x 5 km. The centerpoint of each raster-cell acted as node in the final network
and we used the simulated trajectories to determine all transitions occurring
from one raster-cell to another, as well as the frequency at which those
transitions occurred. This resulted in an edge-list that we translated into a
weighted network using the r-package {\tt igraph} \citep{Gabor.2006}. Because
{\tt igraph} handles edge weights (\(\omega\)) as costs, we inverted the
traversal frequency in each cell by applying \(\omega = \frac{\sum_i^n{Traversal
Frequency_i}/n}{Traversal Frequency_i}\). Finally, we used the weighted network
to calculate the betweenness score of each raster-cell. Betweenness measures how
often a specific raster-cell lies on a shortest path between two other
raster-cells and is therefore a useful metric to detect movement corridors
\citep{BastilleRousseau.2018}.

\subsection{Inter-Patch Connectivity (80\%)}
To exemplify how simulations can be used to provide a network view on
connectivity, we assessed inter-patch connectivity between national parks
located in our study area. The decision to focus on national parks was purely
out of simplicity and does not imply that connections between other regions are
impossible. In fact, the same logic could easily be expanded to include other
protected areas as well. To quantify inter-patch connectivity, we computed the
relative frequency at which dispersers originating from one national park
successfully moved into another national park. Successful movement was said to
be achieved if the individuals' trajectory ever intersected with the
corresponding national park. We also recorded the step-number of the first step
that ever intersected with the national park's polygon. This allowed us to
determine \textit{if} and \textit{how often} dispersers moved between certain
national parks, as well as to estimate \textit{how long} dispersers had to move
to realize those connections.

\section{Results}
\subsection{Movement Model (80\%)}
Compared to the base model reported in \citep{Hofmann.2021}, our most
parsimonious movement model retained several additional interactions between
habitat covariates and movement covariates (\Cref{MovementModel} and
\Cref{MovementModelNumbers}). Although several models received an AIC weight
above zero (Table 1 in Appendix S1), we only considered results from the most
parsimonious model for simplicity. All models with positive AIC weight included
similar covariates (Table S1), so this decision only marginally influenced
subsequent analyses. Plots that aid with the interpretation of the final model
are provided in Appendix S2.

Assuming that all other covariates are held constant, the habitat kernel reveals
that dispersing wild dogs avoid water but prefer its proximity. Similarly,
dispersers avoid areas that are covered by woodlands, yet prefer regions covered
by shrublands or grasslands. Finally, dispersers avoid movement through
landscapes that are dominated by humans. Effect sizes are strong and, except for
effect for \textsf{distance to water}, statistically clear on the 5\%
significance level.

With regards to the movement kernel, the positive estimate for \textsf{cos(ta)}
indicates that dispersers move with directional persistence, unlike what was
proposed by the uniform turning angle distribution. Moreover, directionality is
particularly pronounced when dispersers realize large steps (move quickly), as
indicated by the positive estimates for \textsf{cos(ta):sl} and
\textsf{cos(ta):log(sl)}. Finally, the negative estimate for the interaction
\textsf{sl:LowActivity} reveals that wild dogs realize shorter steps (move
slowly) outside the main activity periods during sunrise and sunset. Aside from
the interaction \textsf{sl:LowActivity}, which appears to strongly influence
movmement behavior, effect sizes are moderate, but mostly significant on the 5\%
significance level.

When looking at the interactions between the movement and habitat kernels, we
observe that movement behavior is contingent on habitat conditions. For example,
there's strong evidence that dispersers realize smaller steps in areas covered
by water or areas covered by wooldand. Conversely, it appears that steps tend to
larger in regions dominated by shrubs/grassland, and shorter when the distance
to water is high. Correspondingly, the model suggests that directionality is
lower in areas dominated by humans, yet more pronounced when dispersers are far
from water. However, except for the effect of \textsf{sl:Water}, effect sizes
are small.

The k-fold cross-validation procedure reveals that our model substantially
outperforms a random guess (\Cref{MovementModel}b). Confidence intervals of
\(\bar{r}_{s, realized}\) and \(\bar{r}_{s, random}\) do not overlap and
therefore proof a reliable prediction. Furthermore, the significant correlation
between ranks and corresponding frequencies for realized steps indicates a good
fit between predictions and observations (\Cref{MovementModel}b). In comparison
to the base model (\(\bar{r}_{s, realized} = -0.55\); \citealp{Hofmann.2021}),
the inclusion of interactions between movement and habitat covariates slightly
improved model performance.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{99_MovementModel}
    \caption{(a) Most parsimonious movement model for dispersing wild dogs. The
    model includes a habitat kernel, a movement kernel, as well as their
    interactions. The horizontal line segments delineate the 90\%, 95\%, and
    99\% Confidence-Intervals for the respective \(\beta\)-coefficients.
    Significance codes: * \(p < 0.10\), ** \(p < 0.05\), *** \(p < 0.01\). (b)
    Results from the k-fold cross validation procedure. The upper plot shows
    rank frequencies of realized steps according to model predictions with known
    preferences, whereas the lower plot shows rank frequencies of realized steps
    when assuming random preferences. The blue ribbon shows the prediction
    interval around a loess smoothing regression that we fitted to ease the
    interpretation of the plots. The significant correlation between rank and
    associated frequency in (b1) highlights that the most parsimonious model
    successfully outperforms a random guess (b2) and assigns comparably high
    selection scores to realized steps.}
    \label{MovementModel}
  \end{center}
\end{figure}

>>>>

%------------------------------------------------------------------------------
%	Titlepage: Header
%------------------------------------------------------------------------------
\title{Dispersal Through Seasonal Landscapes}

@@

%------------------------------------------------------------------------------
%	Abstract
%------------------------------------------------------------------------------
\newpage
\begin{abstract}
This is the abstract.
\end{abstract}

@@

% Taken from Zeller.2020
Connectivity for wildlife is typically modeled in static environments and
estimated for a single snapshot in time (e.g., [6,7]). However, wildlife
responses to landscapes are inherently dynamic, as are the landscapes themselves
(e.g., [8])

Modeling connectivity in a static environment may mask the importance of connectivity for
maintaining wildlife population dynamics and may result in identifying corridors that are only
periodically functional [15,16,18]

\section{Methods}

\section{Results}

\section{Discussion}
